### 2018-05-01 papers/synth/abstract.tex

Vesicle Traffic Systems(VTSs) are the material transport mechanisms
among the compartments inside the biological cells.
%
The compartments are viewed as nodes that are labeled with the
containing chemicals and the transport channels are similarly labeled
edges between the nodes.
%
Understanding VTSs is an ongoing area of research and for many cells
they are partially known.
%
For example, there may be undiscovered edges, nodes, or their labels
in a VTS of a cell.
%
It has been speculated that there are properties that the VTSs must
satisfy.
%
For example, stability, i.e., every chemical that is leaving a
compartment comes back.
%
Many synthesis questions may arise in this scenario, where we want
to complete a partially known VTS under a given property.
%
In the paper, we present novel encodings of the above questions
into the QBF(quantified Boolean formula) satisfiability problems.
%
We have implemented the encodings in a highly configurable tool
and  apply to a couple of found-in-nature VTSs and 
several synthetic graphs.
%
Our results demonstrates that our method can scale upto the
graphs of interest.
%

>>>>

Vesicle Traffic Systems(VTSs) are the material transport mechanisms
among the compartments inside the biological cells.
%
The compartments are viewed as nodes that are labeled with the
containing chemicals and the transport channels
are similarly viewed as labeled edges between the nodes.
%
Understanding VTSs is an ongoing area of research and for many cells
they are partially known.
%
For example, there may be undiscovered edges, nodes, or their labels
in a VTS of a cell.
%
It has been speculated that there are properties that the VTSs must
satisfy.
%
For example, stability, i.e., every chemical that is leaving a
compartment comes back.
%
Many synthesis questions may arise in this scenario, where we want
to complete a partially known VTS under a given property.
%
In the paper, we present novel encodings of the above questions
into the QBF(quantified Boolean formula) satisfiability problems.
%
We have implemented the encodings in a highly configurable tool
and apply to a couple of found-in-nature VTSs and 
several synthetic graphs.
%
Our results demonstrates that our method can scale upto the
graphs of interest.
%

### 2018-05-01 papers/synth/prelim.tex

\paragraph{Activity functions}
We also need to encode that
the activity of the molecules are controlled by activity functions.
%
However, the encoding functions can be given to us many different ways,
for example as a lookup table, or a concise Boolean formula.
%
In the following section, we will assume the appropriate encoding is
used for the functions and represent them by \texttt{NodeFun}$_m$
and \texttt{EdgeFun}$_m$ for node and edge regulations respectively.
%
We will use $\nodef_m$ and $\edgef_m$ to represent functions that
are not defined in VTS.
%
% \begin{align}
% \bigwedge\limits_{i,m} n_{i,m} \limplies a_{i,m} =  \nodef_m (n_{i,1},\dots,n_{i,\mu}) 
% \tag{\texttt{ActiveN}}\label{eq:anb}\\
%    \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies b_{i,j,q,m} = \edgef_m(e_{i,j,q,1}, .., e_{i,j,q,\mu} )
%   \tag{\texttt{ActiveE}}\label{eq:aeb}
% \end{align}
%
Later we will be synthesizing missing activity functions and 
replace $\nodef_m$ and $\edgef_m$ with parameterized constraints that
encode a space of candidate functions.

@@

For synthesis of incomplete systems,
we need properties properties against which we synthesize the missing parts.
%
Here we will discussed two such properties proposed in earlier
works~\cite{SASBpaper}.
%

@@

\paragraph{$k$-connected property}
%
$k$-connectedness is a robustness property.
%
Let us use $d_{i,j,q}$ to indicate $q$th edge between $i$ and $j$ is dropped
and $r'_{i,j}$ to indicate if there is a path from $i$ to $j$ in
the modified VTS.
%
% To check whether $k$-connected is a
% necessary condition, we remove (drop) $k-1$ edges from the graph and
% if it disconnects the graph, and we get an assignment.
% %
% We have a graph that is not $k$-connected.
%
In the following, $\texttt{Drop}(d,k)$  encode that only
existing edges can be dropped and exactly $k-1$ edges are dropped.
%
$\texttt{Reach}(d,r')$ defines reachability in the modified VTS,
therefore we use a new variable $r'_{i,j}$ to encode reachability from
$i$ to $j$.
%
$\texttt{Connected}(r')$ says that all nodes are reachable from any
other node.
% In the following constraint, we encode $r'_{i,j}$ is true if there is an
% edge $(i,\_,j)$ and it is not dropped, or there is a node
% $i^{\prime}$ such that, there is an edge $(i,\_,i^{\prime})$ which is
% not dropped and $r'_{i',j}$ is true.
\begin{align*}
  \texttt{Drop}(d,k) = & 
  \bigwedge\limits_{i,j,q} d_{i,j,q} \limplies e_{i,j,q}  \land 
  \sum_{i,j,q} d_{i,j,q} = k-1\\
  \texttt{ReachAbove}(d,r') = &
  \bigwedge\limits_{i,j}  [\bigvee_{q} (e_{i,j,q} \land  \neg d_{i,j,q}) \lor  (\bigvee_{i' \neq i}  r^{\prime}_{i',j} \land  \bigvee_{q} (e_{i,i',q} \land \neg d_{i,i',q}) ] \limplies r^{\prime}_{i,j}\\
  \texttt{ReachBelow}(d,r') = &
   \bigwedge\limits_{i,j}  r^{\prime}_{i,j} \limplies [\bigvee_{q} (e_{i,j,q} \land  \neg d_{i,j,q}) \lor  (\bigvee_{i' \neq i}  r^{\prime}_{i',j} \land  \bigvee_{q} (e_{i,i',q} \land \neg d_{i,i',q}) ]\\
  \texttt{Connected}(r') = & \Land\limits_{i,j} (r^{\prime}_{i,j} \lor r^{\prime}_{j,i})
\end{align*}
We will be synthesizing for both connected any disconnected graphs.
%
We also define the following macros. 
\begin{align*}
  \texttt{Connected}(k) = & \forall d.\;
          (\texttt{Drop}(d,k) \limplies \exists r'.\;\texttt{ReachBelow}(d,r')
                                \land \texttt{Connected}(r'))\\
  \texttt{Disconnected}(k) = & \exists d.\;
          (\texttt{Drop}(d,k) \land \exists r'.\;\texttt{ReachAbove}(d,r')
                                \land \lnot \texttt{Connected}(r'))
\end{align*}
Since the free variables in $\texttt{Connectedness}(k)$ are universally
quantified, please note that the connectedness property introduces the two
quantifier alternations.
%

>>>>

\paragraph{Activity functions}
We also need to encode that
the activity of the molecules are controlled by activity functions.
%
However, the encoding functions can be given to us many different ways,
for example as a lookup table, or a concise Boolean formula.
%
In the following section, we will assume the appropriate encoding is
used for the functions and represent them by \texttt{NodeFun}$_m$
and \texttt{EdgeFun}$_m$ for node and edge regulations respectively.
%
We will use $\nodef_m$ and $\edgef_m$ to represent functions that
are not defined in a VTS.
%
% \begin{align}
% \bigwedge\limits_{i,m} n_{i,m} \limplies a_{i,m} =  \nodef_m (n_{i,1},\dots,n_{i,\mu}) 
% \tag{\texttt{ActiveN}}\label{eq:anb}\\
%    \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies b_{i,j,q,m} = \edgef_m(e_{i,j,q,1}, .., e_{i,j,q,\mu} )
%   \tag{\texttt{ActiveE}}\label{eq:aeb}
% \end{align}
%
Later we will be synthesizing missing activity functions and 
replace $\nodef_m$ and $\edgef_m$ with parameterized constraints that
encode a space of candidate functions.

@@

For synthesis of incomplete systems,
we need properties  against which we synthesize the missing parts.
%
Here we will discussed two such properties proposed in earlier
works~\cite{SASBpaper}, namely stability and $k$-connectedness.
%

@@

\paragraph{$k$-connected property}
%
$k$-connectedness expresses robustness against failure of few edges.
%
Let us use $d_{i,j,q}$ to indicate $q$th edge between $i$ and $j$ is failed
and $r'_{i,j}$ to indicate if there is a path from $i$ to $j$ in
the modified VTS.
%
% To check whether $k$-connected is a
% necessary condition, we remove (drop) $k-1$ edges from the graph and
% if it disconnects the graph, and we get an assignment.
% %
% We have a graph that is not $k$-connected.
%
In the following, $\texttt{Drop}(d,k)$  encode that only
existing edges can be failed and exactly $k-1$ edges are failed.
%
Both $\texttt{ReachBelow}(d,r')$ and $\texttt{ReachAbove}(d,r')$ defines
reachability in the modified VTS,
therefore we use a new variable $r'_{i,j}$ to encode reachability from
$i$ to $j$.
%
$\texttt{Connected}(r')$ says that all nodes are reachable from any
other node.
% In the following constraint, we encode $r'_{i,j}$ is true if there is an
% edge $(i,\_,j)$ and it is not dropped, or there is a node
% $i^{\prime}$ such that, there is an edge $(i,\_,i^{\prime})$ which is
% not dropped and $r'_{i',j}$ is true.
\begin{align*}
  \texttt{Drop}(d,k) = & 
  \bigwedge\limits_{i,j,q} d_{i,j,q} \limplies e_{i,j,q}  \land 
  \sum_{i,j,q} d_{i,j,q} = k-1\\
  \texttt{ReachAbove}(d,r') = &
  \bigwedge\limits_{i,j}  [\bigvee_{q} (e_{i,j,q} \land  \neg d_{i,j,q}) \lor  (\bigvee_{i' \neq i}  r^{\prime}_{i',j} \land  \bigvee_{q} (e_{i,i',q} \land \neg d_{i,i',q}) ] \limplies r^{\prime}_{i,j}\\
  \texttt{ReachBelow}(d,r') = &
   \bigwedge\limits_{i,j}  r^{\prime}_{i,j} \limplies [\bigvee_{q} (e_{i,j,q} \land  \neg d_{i,j,q}) \lor  (\bigvee_{i' \neq i}  r^{\prime}_{i',j} \land  \bigvee_{q} (e_{i,i',q} \land \neg d_{i,i',q}) ]\\
  \texttt{Connected}(r') = & \Land\limits_{i,j} (r^{\prime}_{i,j} \lor r^{\prime}_{j,i})
\end{align*}
We will be synthesizing for both connected and disconnected graphs.
%
We define the following macros. 
\begin{align*}
  \texttt{Connected}(k) = & \forall d.\;
          (\texttt{Drop}(d,k) \limplies \exists r'.\;\texttt{ReachBelow}(d,r')
                                \land \texttt{Connected}(r'))\\
  \texttt{Disconnected}(k) = & \exists d.\;
          (\texttt{Drop}(d,k) \land \exists r'.\;\texttt{ReachAbove}(d,r')
                                \land \lnot \texttt{Connected}(r'))
\end{align*}
Since the free variables in $\texttt{Connectedness}(k)$ are universally
quantified, the connectedness property introduces the two
quantifier alternations.
%
